---
title: "Exercise - some linear models"
author: "Søren Højsgaard"
format: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<---SOLUTION
some text
--->



We will use the `palmerpenguins` data set. It is a data set with
information on penguins. We will use the `body_mass_g` as the response
variable and `species`, `island`, `sex`, `culmen_length_mm`,
`culmen_depth_mm`, `flipper_length_mm`, and `year` as explanatory
variables.

```{r}
dat <- palmerpenguins::penguins
dat <- dat[complete.cases(dat),]
head(dat)
```

```{r}
library(ggplot2) 
library(broom)
```



We ask the following question: 

1. How well can we predict the body mass of a penguin based on species and flipper length?

2. Are predictions improved by including other variables?

We encourage you to take inspiration from the lecture slides/notes.

```{r}
pl0 <- dat |> ggplot(aes(x=flipper_length_mm, y=body_mass_g, color=species)) + geom_jitter(width=0.1)
pl0 + geom_smooth(se=FALSE)
```

```{r}
mm1 <- lm(body_mass_g ~ flipper_length_mm, data = dat)
mm2 <- lm(body_mass_g ~ species, data = dat)
mm3 <- lm(body_mass_g ~ species + flipper_length_mm, data = dat)
mm4 <- lm(body_mass_g ~ species * flipper_length_mm, data = dat)
```

```{r}
## SOLUTION
pl0 + geom_line(aes(y=fitted(mm1)))
mm1 |> tidy()
```

```{r}
## SOLUTION
pl0 + geom_line(aes(y=fitted(mm2)))
mm2 |> tidy()
```

```{r}
## SOLUTION
pl0 + geom_line(aes(y=fitted(mm3)))
mm3 |> tidy()
```

```{r}
## SOLUTION
pl0 + geom_line(aes(y=fitted(mm4)))
mm4 |> tidy()
```


```{r}
## SOLUTION
ml <- list(mm1, mm2, mm3, mm4)
i <- sample(nrow(dat), size=0.8*nrow(dat))
i |> head()
train <- dat[i,]
test <- dat[-i,]
```


```{r}
## SOLUTION
sapply(ml, function(x) modelr::rmse(x, dat))
ml_train <- sapply(ml, function(x) update(x, data=train))
sapply(ml_train, function(x) modelr::rmse(x, train))
sapply(ml_train, function(x) modelr::rmse(x, test))
```



```{r}
## YOUR CODE
## here is a hint
x <- 1:10
```


```{r}
## YOUR CODE
```

<---SOLUTION
some text
--->
